---
title: "Transform Data Pt. 2"
date: "2019-10-1"
---

```{r setup}
library(tidyverse)
#install.packages("babynames")
library(babynames)
babynames
```

# dplyr
```{r line graph}
ggplot(data=babynames)+
  geom_line(mapping=aes(x=year, y=prop))
```


## Your Turn 1

Alter the code to select just the `n` column:

```{r}
select(babynames, n)
```

## Quiz

Which of these is NOT a way to select the `name` and `n` columns together?

```{r}
select(babynames, -c(year, sex, prop))
select(babynames, name:n)
select(babynames, starts_with("n"))
#select(babynames, ends_with("n"))
```

## Your Turn 2

Use `filter`, `babynames`, and the logical operators to find:

* All of the names where prop is greater than or equal to 0.08  
* All of the children named "Sea"  

```{r}
filter(babynames, prop>=0.08)
filter(babynames, name=="Sea")
filter(babynames, name=="Garrett", year==1880  )
```

## Your Turn 3

Use Boolean operators to return only the rows that contain:

* _Boys_ named Sue  
* Names that were used by exactly 5 or 6 children in 1880  
* Names that are one of Acura, Lexus, or Yugo

```{r}
filter(babynames, name=="Sue", sex=="M")
filter(babynames, year==1880 & (n==5|n==6))
filter(babynames, name=="Acura"| name=="Lexus"| name=="Yugo")
```

## Help Me

What is the smallest value of n?
What is the largest?

```{r}
arrange(babynames, n)
arrange(babynames, desc(n))
```

## Your Turn 4

Use `%>%` to write a sequence of functions that: 

1. Filters babynames to just the girls that were born in 2017, *then...*  
2. Selects the `name` and `n` columns, *then...*  
3. Arranges the results so that the most popular names are near the top.

```{r}
girls_2017<-filter(babynames, sex=="F", year==2017)
girls_2017<-select(girls_2017, name, n, prop)
girls_2017<-arrange(girls_2017, desc(n), prop)
girls_2017

babynames %>%
  filter(year==2017, sex=="F") %>%
  select(name, n, prop) %>%
  arrange(desc(n), prop)
``` 

## Your Turn 5 - Exam

1. Trim `babynames` to just the rows that contain **your** `name` and **your** `sex`  
2. Trim the result to just the columns that will appear in your graph (not strictly necessary, but useful practice)  
3. Plot the results as a line graph with `year` on the x axis and `prop` on the y axis

```{r}
babynames %>%
  filter(name=="Armando", sex=="M") %>%
  select(name, sex, year, prop)%>%
  arrange(year) %>%
  ggplot() +
  geom_line(aes(year, prop))
```

## Your Turn 6

Copmplete the code below to extract the rows where `name == "Khaleesi"`. Then use `summarise()` and `sum()` and `min()` to find:

1. The total number of children named Khaleesi
2. The first year Khaleesi appeared in the data

*(Hint: Be sure to remove each `_` before running the code)*

```{r}
babynames %>% 
  filter(name=="Khalessi") %>%
  summarise(total = sum(n), first = min(year))
```

## Your Turn 7

Use `group_by()`, `summarise()`, and `arrange()` to display the ten most popular names. Compute popularity as the *total* number of children of a single gender given a name.

*(Hint: Be sure to remove each `_` before running the code)*

```{r}
babynames %>%
  group_by(name, sex) %>% 
  summarise(total = sum(n))%>%
  arrange(desc(total)) %>% 

  ungroup() %>% 
  slice(1:10) %>% 
  ggplot() +
    geom_col(mapping = aes(x = fct_reorder(name, 
      desc(total)), y = total, fill = sex)) +
    labs(x = "name")

```


## Your Turn 8

Fill in the blanks to: 
Rank each name _within its year and sex_. 

*(Hint: Be sure to remove each `_` before running the code)*

```{r}
babynames %>% 
  group_by(year, sex) %>% 
  mutate(rank = min_rank(desc(c(prop))))
```


***

# Take aways

* Extract variables with `select()`  
* Extract cases with `filter()`  
* Arrange cases, with `arrange()`  

* Make tables of summaries with `summarise()`  
* Make new variables, with `mutate()`  
* Do groupwise operations with `group_by()`

* Connect operations with `%>%`  


